---
title: "Map of participants in Open Science for Synthesis"
author: "A.J. Perez-Luque, S. Varela, T. Assal"
date: "Sep 2014"
output:
  md_document:
    variant: markdown_github  

  
---

In this document we create a map of the origin of the participants of the Open Science for Synthesis 2014 hold at National Center for Ecological Analysis and Synthesis ([NCEAS](https://www.nceas.ucsb.edu/)) and University of North Carolina???s Renaissance Computing Institute ([RENCI](http://www.renci.org/)). 

We created a map from the list of participants provided by NCEAS (available [here](https://www.nceas.ucsb.edu/news/open-science-synthesis-announcing-2014-program-participants)) and RENCI (available [here](http://renci.org/news/participants-from-around-the-world-to-participate-in-bicoastal-open-science-for-synthesis-course/)). 

## Geolocations
In a first part we prepare the data and get the coordinates of the cities. To do this we used the `geocode` function of `ggmap` [package](http://cran.r-project.org/web/packages/ggmap/index.html).  

```{r ReadData}
# set directory 
di  <- '/Users/ajpelu/Dropbox/MS/nota_ECOSISTEMAS'

# load library
library('ggmap')
library('ggplot2')
library('grid')

# read table 
mydata <- read.csv(paste(di,'/data/participants.csv', sep=''),header=T, sep=';')
```

```{r GetGeolocation, message=FALSE, warning=FALSE}
# Create a name with city and Country 
mydata$location <- paste(mydata$city, mydata$country, sep=', ')

# Get coordinates from Google Maps API (see geocode function)
mydata$long <- geocode(mydata$location)[,1]
mydata$lat <- geocode(mydata$location)[,2]
```

Now we aggregate the data by city and count the number of participants by city. We hold the name of the center in which the participant attended to the course. 
```{r Aggregate}
# Aggregate by city
mycities <- aggregate(mydata$city, by=list(mydata$city, mydata$country, mydata$lat, 
                                           mydata$long, mydata$center),
                      FUN = length)
names(mycities) <- c('city','country','lat','long','center','count')
mycities
```

## Plot the map 
```{r mapPlot, fig.width=8, fig.height=5, cache=TRUE}
# colour world map 
micolor <- 'gray80'
color.nceas <- '#41ab5d'
color.renci <- '#d7301f'

# Get a world map
wm <- map_data('world')

p <- ggplot() + geom_polygon(data=wm, aes(x=long, y=lat, group=group), fill=micolor)
p1 <- p + geom_point(data=mycities,aes(x=long, y=lat, size=factor(count), shape=center, color=center)) + 
  scale_shape_manual(name='Center', values=c(15,16))+ 
  scale_size_manual(values=c(3,4,5), guide=FALSE)+ 
  scale_colour_manual(name='Center', values=c(color.nceas,color.renci))+
  theme_bw() + theme(panel.grid.major=element_blank(), 
                     panel.grid.minor.y = element_blank(),
                     legend.key.size = unit(1.2, "cm"),
                     legend.key=element_blank())+
  guides(colour = guide_legend(override.aes = list(size=5))) # To increase the size of points in legend

# Render the map
p1
```

